Dark matter halo shapes at $z=0$ in the Auriga simulations

### 2019-08-03 Document/paper.tex
edit

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_DM.pdf}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_MHD.pdf}
\end{center}
\caption{Axial ratios for all simulated halos.
  Left/right panel corresponds to DMO/MHD simulations, respectively.
  Triangles (squares) represent the measurements at $R_{200}/16$
  ($R_{200}$) that correspond to physical distances of $\approx14$ kpc
  ($\approx 230$ kpc).
  Here we visualize three main population trends.
  First, in DMO simulations halos are rounder in the outskirts
  than in the inner part.
  Second, halos in MHD are rounder in the inner regions than in
  the outskirts (opposite to the DMO trend). 
  Third, halos in MHD are rounder than its DMO counterparts.
  The first and second points are claryfied in Figure
  \ref{fig:triaxial_cumulative} through the cumulative triaxility
  distributions, 
  while the third trend is shown in Figure
  \ref{fig:delta_triaxial_cumulative} through the cumulative
  distribution in the triaxility changes between MHD and DMO at all
  radii.}
  \label{fig:triaxiality_plane}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.8\columnwidth]{triaxialiy_distro_DM.pdf}
\includegraphics[width=0.8\columnwidth]{triaxialiy_distro_MHD.pdf}
\caption{Triaxility cumulative distribution at five different radii.
  Right/left panel corresponds to DMO/MHD simulations, respectively. 
  In DMO simulations the median triaxiality at all radii is larger
  than $2/3$. Furthermore, the triaxility increases as one moves
  towards the inner part of the halo.
  In MHD simulations these trend reverses.
  The median triaxility at all radii is smaller than $1/3$ and the
  halo is less triaxial as on moves towards the stellar disc.}
\label{fig:triaxial_cumulative}
\end{figure*}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{delta_triaxiality_distro.pdf}
\caption{
  Cumulative distribution for the changes in triaxility in simulations
  with different physics, $\Delta T=T_{\rm MHD}-T_{\rm DMO}$.
  At the virial radius all halos become less triaxial in the MHD
  simulations, the change is stronger towards the halo center.
  Only two haloes in the sample show the opposite trend of becoming
  more triaxial (but only at the smaller radii).} 
\label{fig:delta_triaxial_cumulative}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{angles_alignment_DM.pdf}
\includegraphics[width=1.0\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{\textbf{RP: I think the upper row that compares the shape of the halo from the DMO sim with the orientation of the disk in the MHD sim is hard to interpret. Maybe instead show in the first row the angle against a fixed axis (e.g. total angular momentum of the halo?)} Angles between the principal axis of the dark matter halo
  shape and the angular momentum of the stellar disc 
  as a function of the radius at
  which the halo shape directions are measured.
  Each panel compares the alignment of the corresponding
  major/middle/minor axis in the halo.
  Thin lines correspond to each one of the thirty halos in the sample
  while the thick line traces the median value at every radius.
  In the upper row the haloes come from the DMO simulation, 
  showing that the directions in the ellipsoids describing the shape
  are constant as a function of radius for the most part of the sample.
  In the lower row the haloes come from the MHD simulation providing a 
  self-consistent comparison with the stellar discs. 
  In this case the dark matter shells twists significantly in $1/5$ of the sample. For the remaining haloes the alignment is almost constant as a function of radius.
}
\label{fig:cumulative_alignment}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{correlation_T_MHD_disc.pdf}
\end{center}
\caption{Correlations between the halo triaxility at different radii
  and baryonic disc properties. 
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient (mean value and uncertainty estimated via
  jackknife resampling).
  The line is the best minimum squares fit to a line.
  The x-axis in the first column is the gas density at the center of
  the galaxy with in a sphere of radius  $1$ kpc \citep{Pakmor17};
  the second column shows the disc to total mass ratio and the last
  column includes the disc optical radius defined to be the radius at which the
  $B$-band surface brightness drops below 25 mag arcsec$^{-2}$ \citep{auriga}.
  The largest correlations are found for the two smaller radii
  ($0.06R_{200}$ and $0.12R_{200}$).
  Extended and massive stellar discs with a low gas content at its
  core are correlated with low dark matter triaxilities.
  The correlation decreases as one approaches larger radii.}
\label{fig:disc_correlations}
\end{figure*}

@@

{\bf Recently \cite{2017MNRAS.469..594B} reported that MW-like DM haloes can
be roughly splint into two distinct families: stalled and accreting.
Stalled haloes assembled early and have remained almost unchanged
in the last $7$ Gyr, while accreting halos have been growing
constantly.
\cite{2017MNRAS.469..594B} also link this dicotomy to the dark matter
filaments feeding the halos. 
Stalled halos at $z=0$ are embbeded in a single large thick filament, while
accreting halos are pierced by many thinner filaments. 
To what extent the twisting and high triaxility present in the MHD
halos can be linked to this dicotomy?
In Figure \ref{fig:twisting_age} we show the correlation between the
twisting and the age of the bulge, the age of the stellar disc and the
DM halo formation redshift. 
The ages are estimated directly as the mean value in the stellar age
distribution for each component, while the formation redshift is
defined as the redshif at which the halo already has half of its $z=0$
mass. 
In the three cases we find a correlation between ages and twisting
that goes in the direction claimed by \cite{2017MNRAS.469..594B}:
components with early assembly (i.e. stalled) have a smaller degree of
twisting.
A detailed study with respect to the cosmic web location is beyond the
scope of our paper.}

@@

In this paper we measured the shape of 30 isolated Milky Way sized
dark matter haloes simulated in the Auriga project using the zoom-in
technique. 
The halos were with two different setups:
dark matter only (DMO) simulation and full magnetohydrodynamics (MHD)
including star formation and feedback.
We used the shape measurement algorithm by \cite{Allgood06} on the
dark matter halos of these simulations to quantify the halo shape as a
function of radius and the degree of aligment between the stellar disc
and the halo.

We find that MHD halos are rounder than DMO halos at every radius.
MHD halos tend towards more oblate, sometimes almost spherical, shapes
($T<1/3$), while DMO halos tend towards more prolate shapes ($T>2/3$).  
The rounding effect is more noticeable as one moves closer to galactic
disc and strongly  correlates with baryonic properties of the disc
More precisely, the triaxility is smaller for extended and massive
stellar discs with low gas densities at its core.

We also measured the alignment of the halo with the stellar disc's angular momentum at different radii.
For the majority of the sample the angular momentum is strongly aligned with the minor axis of the halo at all measured radii.
However, in some halos the alignment changes noticeably with radius. 
This alignment evolution implies a radial twisting between the ellipsoids
describing the halo shape. 
We quantify this twist with the standard deviation of the angle
between the angular momentum and the halo minor axis at radii below $\leq 0.5R_{200}$ and find that higher gas densities correlate with larger twisting values.

>>>>

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_DM.pdf}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_MHD.pdf}
\end{center}
\caption{Axial ratios for all simulated halos.
  Left/right panel corresponds to DMO/MHD simulations, respectively.
  Triangles (squares) represent the measurements at $R_{200}/16$
  ($R_{200}$) that correspond to physical distances of $\approx14$ kpc
  ($\approx 230$ kpc).
  Here we visualize three main population trends.
  First, in DMO simulations halos are rounder in the outskirts
  than in the inner part.
  Second, halos in MHD are rounder in the inner regions than in
  the outskirts (opposite to the DMO trend). 
  Third, halos in MHD are rounder than its DMO counterparts.
  These three points are further clarified in Figure 
  The first and second points are claryfied in Figure
  \ref{fig:triaxial_cumulative} through cumulative triaxility
  distributions}
  \label{fig:triaxiality_plane}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.32\textwidth]{triaxialiy_distro_DM.pdf}
\includegraphics[width=0.32\textwidth]{triaxialiy_distro_MHD.pdf}
\includegraphics[width=0.32\textwidth]{delta_triaxiality_distro.pdf}
\caption{Triaxility cumulative distribution at five different radii.
  Right/middle panel corresponds to the cumulative distribution of
  triaxiality in DMO/MHD simulations, respectively. 
  In DMO simulations the median triaxiality at all radii is larger
  than $2/3$. Furthermore, the triaxility increases as one moves
  towards the inner part of the halo.
  In MHD simulations these trend reverses.
  The median triaxility at all radii is smaller than $1/3$ and the
  halo is less triaxial as on moves towards the stellar disc.
  The right panel, shows the cumulative distribution for the changes
  in triaxility in simulations with different physics, $\Delta
  T=T_{\rm MHD}-T_{\rm DMO}$. 
  At the virial radius all halos become less triaxial in the MHD
  simulations, the change is stronger towards the halo center.
  Only two haloes in the sample show the opposite trend of becoming
  more triaxial (but only at the smaller radii).} 
\label{fig:triaxial_cumulative}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{angles_alignment_DM.pdf}
\includegraphics[width=0.9\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{\textbf{RP: I think the upper row that compares the shape of the halo from the DMO sim with the orientation of the disk in the MHD sim is hard to interpret. Maybe instead show in the first row the angle against a fixed axis (e.g. total angular momentum of the halo?)} Angles between the principal axis of the dark matter halo
  shape and the angular momentum of the stellar disc 
  as a function of the radius at
  which the halo shape directions are measured.
  Each panel compares the alignment of the corresponding
  major/middle/minor axis in the halo.
  Thin lines correspond to each one of the thirty halos in the sample
  while the thick line traces the median value at every radius.
  In the upper row the haloes come from the DMO simulation.
  The first thing to notice is that the major axis is statistically
  perpendicular to the angular momentum of the disc that will form in
  the MHD simulations. The other two directions do not show any
  correlation. 
  The second important result is that the directions in the ellipsoids
  describing the shape are constant as a function of radius except two
  cases. 
  In the lower row the haloes come from the MHD simulation providing a 
  self-consistent comparison with the stellar discs. 
  In this case the sample presents a good alignment of the angular
  momentum with the minor halo axis. 
  Furthermore, the dark matter shells twists significantly in $1/5$ of
  the sample. For the remaining haloes the alignment is almost
  constant as a function of radius. 
}
\label{fig:cumulative_alignment}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{correlation_T_MHD_disc.pdf}
\end{center}
\caption{Correlations between the halo triaxility at different radii
  and baryonic disc properties. 
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient (mean value and uncertainty estimated via
  jackknife resampling).
  The line is the best minimum squares fit to a line.
  The x-axis in the first column is the gas density at the center of
  the galaxy with in a sphere of radius  $1$ kpc \citep{Pakmor17};
  the second column shows the disc to total mass ratio and the last
  column includes the disc optical radius defined to be the radius at which the
  $B$-band surface brightness drops below 25 mag arcsec$^{-2}$ \citep{auriga}.
  The largest correlations are found for the two smaller radii
  ($R_{200}/16$ and $R_{200}/8$).
  Extended and massive stellar discs with a low gas content at its
  core are correlated with low dark matter triaxilities.
  The correlation decreases as one approaches larger radii.}
\label{fig:disc_correlations}
\end{figure*}

@@

{\bf Recently \cite{2017MNRAS.469..594B} reported that MW-like DM haloes can
be roughly splint into two distinct families: stalled and accreting.
Stalled haloes assembled early and have remained almost unchanged
in the last $7$ Gyr, while accreting halos have been growing
constantly.
\cite{2017MNRAS.469..594B} also link the two populations to
differences in the dark matter filaments feeding the halos. 
Stalled halos at $z=0$ are embbeded in a single large thick filament, while
accreting halos are pierced by many thinner filaments. 
To what extent the twisting and high triaxility present in the MHD
halos can be linked to this dicotomy?
In Figure \ref{fig:twisting_age} we show the correlation between the
twisting and the age of the bulge, the age of the stellar disc and the
DM halo formation redshift. 
The ages are estimated directly as the mean value in the stellar age
distribution for each component, while the formation redshift is
defined as the redshif at which the halo already has half of its $z=0$
mass. 
In the three cases we find a correlation between ages and twisting
that goes in the direction claimed by \cite{2017MNRAS.469..594B}:
components with early assembly (i.e. stalled) have a smaller degree of
twisting.
A detailed study with respect to the cosmic web location is beyond the
scope of our paper.}

@@

In this paper we measured the shape of 30 isolated Milky Way sized
dark matter haloes simulated in the Auriga project using the zoom-in
technique. 
The halos were simulated in two different setups:
dark matter only (DMO) simulation and full magnetohydrodynamics (MHD)
including star formation and feedback.
We used the shape measurement algorithm by \cite{Allgood06} on the
dark matter halos of these simulations to quantify the halo shape as a
function of radius and the degree of aligment between the angular
momentum in stellar disc and the dark matter halo.

We find that MHD halos are rounder than DMO halos at every radius.
MHD halos tend towards more oblate, sometimes almost spherical, shapes
($T<1/3$), while DMO halos tend towards more prolate shapes ($T>2/3$).  
The rounding effect is more noticeable as one moves closer to the galactic
disc and strongly  correlates with baryonic properties of the disc
More precisely, the triaxility is smaller for extended and massive
stellar discs with low gas densities at its core.

We also measured the alignment of the halo with the stellar disc's angular momentum at different radii.
For the majority of the sample the angular momentum is strongly aligned with the minor axis of the halo at all measured radii.
However, in some halos the alignment changes noticeably with radius. 
This alignment evolution implies a radial twisting between the ellipsoids
describing the halo shape. 
We quantify this twist with the standard deviation of the angle
between the angular momentum and the halo minor axis at radii below
$\leq 0.5R_{200}$ and find that younger bulges and higher gas
densities correlate with larger twisting values.
